### 2022-03-17 paper/manuscript.Rmd
checked manuscript and replied to comments

$$
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
$$
The resulting parameter estimates are those that give the best predictions of the dependent variable in the present dataset.
Penalized regression, by contrast, adds a penalty term to the RSS.
One commonly used penalty is the L1-norm of the regression coefficients, or LASSO penalty ([@hastieElementsStatisticalLearning2009]),
which corresponds to the sum of their absolute values.
This gives the penalized residual sum of squares:

@@

The standard linear model estimates an intercept, which reflects the expected value of the outcome when all predictors are equal to zero, and regression coefficients for the effect of moderators.
In some cases, it may be desirable to omit the intercept.
For example, if an analysis contains categorical predictors,
these can be encoded as dummy variables, with values $x \in \{0, 1\}$.
For a variable with $c$ categories, the number of dummy variables must be equal to $c-1$;
the omitted category functions as a reference category, and its expected value is represented by the model intercept $b_0$.
This so-called *regression specification* of a model may be useful when there is a meaningful reference category.
For example, imagine a study on the effectiveness of interventions for specific phobia with two types of interventions: An industry-standard intervention, and a more recently developed one.
In this case, it might make sense to code the industry standard intervention as the reference category,
and dummy-code the new contender.
The model will then estimate whether the new contended has an effect size significantly lower or higher than the industry standard.
In other cases, there may not be a straightforward reference category.
For example, imagine a study on the effectiveness of one intervention for specific phobia in two continents.
In such cases, the average effect in both continents may be estimated by omitting the intercept, and including all $c$ dummy variables.
This so-called *ANOVA specification* of a model estimates a mean for all dummy-coded categories.
In BRMA, as in other R functions, one can use ANOVA specification by explicitly removing the intercept from the model formula; for example, if `yi` is the effect size and $C$ a categorical moderator, regression specification with $c-1$ dummies is specified as `yi ~ C`, and ANOVA specification with $c$ dummies is specified as `yi ~ -1 + C`.
The appropriate standardization of dummy variables in penalized regression analyses is a topic of ongoing debate.
Suffice it to say that, for greater interpretability, users may want the freedom to leave dummy variables unstandardized.
This is possible in `brma()` by ...

@@

Many different prior distributions exist that have specific properties that enable them to shrink small regression coefficients towards zero, while keeping substantial coefficients large ([@vanErpOberskiMulder2019]). Some prior distributions have been shown to result in exactly the same solutions as classical, frequentist penalties, such as the LASSO. Specifically, placing independent Laplace (i.e., double exponential) priors on the regression coefficients results in posterior modes that are equal to the lasso estimates ([@ParkCasella2008]). In addition, many prior distributions have been developed and investigated specifically for the purpose of providing good shrinkage properties, meaning that the prior pulls small regression coefficients towards zero, without exerting any influence on substantial regression coefficients. A popular prior in this regard is the horseshoe prior ([@CarvalhoPolsonScott2010]). An advantage of the horseshoe prior compared to the lasso is the fact that it has heavier tails. As a result, it will not shrink and therefore bias substantial coefficients as much as the LASSO prior with its lighter tails.

@@

Bayesian penalized meta-analysis is implemented in the function `brma()` in the R-package `pema`.
For estimation, it depends on Stan, a probabilistic programming language that uses Hamiltonian Monte Carlo to sample from the posterior distribution ([@StanManual]).
Being written in C++, Stan is computationally efficient, but models must be compiled prior to estimation.
This results in substantial computational overhead.
To avoid this overhead, `pema` uses pre-compiled models corresponding to random-effects and three-level meta-regression, with and without an intercept.
Future updates may bring additional models.
At the time of writing, `brma()` supports two priors: the LASSO and the regularized horseshoe.
The LASSO prior is implemented as follows:
<!-- CJ: Sara, kan jij checken of La de juiste notatie is voor de Laplace prior? -->

$$
\beta_{j} \sim \text{La}(0, \frac{s}{\lambda})
$$
where $s$ is a global scale parameter, and $\lambda$ is an inverse-tuning parameter.
By default in `brma()`, the global scale parameter is set to 1,
and the inverse-tuning parameter is given a $\chi^2$ prior with 1 degree of freedom.
Its value is thus optimized during model estimation.

@@

$$
\begin{aligned}
\beta_{j} &\sim N(0, \tilde{\tau}_j^2\lambda) \text{, with } \tilde{\tau}_j^2 = \frac{c^2 \tau_j^2}{c^2 + \lambda^2 \tau_j^2}\\
\lambda &\sim \text{half-t}(\nu_1, 0, \lambda^2_0)\\
\tau_j &\sim \text{half-t}(\nu_2, 0, 1)\\
c^2 &\sim \Gamma^{-1}(\frac{\nu_3}{2}, \frac{\nu_3 s^2}{2})
\end{aligned}
$$
This extension of the horseshoe is more numerically stable in certain cases.
In addition, it allows the user to explicitly include prior information regarding the expected number of relevant moderators by setting the argument par_ratio to the ratio of the expected number of non-zero coefficients to the expected number of zero coefficients. <!-- CJ: Je spreekt hier over het argument par_ratio van de functie in Stan. Maar hier in de tekst verwacht ik uitleg over de formule. Zien we par_ratio terug in die formule?-->
In this formula, $\lambda^2_0$ is a global scale parameter that affects the overall shrinkage of the prior, with smaller values resulting in more shrinkage. The default value in `brma()` is 1. 
The thickness of the tails is controlled by two degrees of freedom parameters, $\nu_1$ and $\nu_2$, which default to 1 in `brma()`.
Increasing these degrees of freedom parameters results in a prior with lighter tails,
which is strictly no longer a horseshoe prior.
However, in certain cases<!-- Sara, wat voor cases?--> these lighter tails can aid model convergence. 
The regularized horseshoe differs from the standard horseshoe in the specification of a finite "slab". This slab ensures at least some regularization of large coefficients and as a consequence,
more stable results.
This slab is governed by a degrees of freedom parameter ($\nu_3$, set to 4) and a scale parameter ($s$, set to 1). <!--Sara, eigenlijk is het niet helemaal duidelijk wat al deze nummers doen. Kunnen we een interactieve plot maken waarmee gebruikers kunnen zien hoe aanpassingen de vorm van de prior beinvloeden?-->

@@

Predictive performance reflects how well the algorithm is able to predict data not used to estimate the model parameters, in other words, it indicates the generalizability of the model.
To compute it, for each iteration of the simulation both a training dataset and a testing dataset are generated.
The model is estimated on the training data, which has a varying number of cases according to the simulation conditions.
Predictive performance is then operationalized as the explained variance in the testing data, $R^2_{test}$.
The testing data has 100 cases in all simulation conditions.
The $R^2_{test}$ reflects the fraction of variance in the testing data explained by the model,
relative to the variance explained by the mean of the training data.
For a predictive performance measure, it is necessary to use the mean of the training data, as the mean of the testing set is a descriptive statistic of that sample. **ik vind deze zin niet heel duidelijk**
The resulting metric $R_{test}^{2}$ is expressed by the following equation:

@@

To examine performance in a range of realistic meta-analysis scenarios,
several design factors were manipulated:
The number of studies in the training data $k \in (20, 40, 100)$, the average within-study sample size $\bar{n} \in (40, 80, 160)$, the population effect size of relevant moderators $\beta \in (0, .2, .5, .8)$, the number of moderators $p \in (2, 3, 6)$, and residual heterogeneity $\tau^{2} \in (.01, .04, .1)$. **het aantal moderatoren is redelijk laag; ik zou vooral nog voordelen van brma verwachten met meer moderatoren maar is dat realistisch in de praktijk? Zo ja, dan is het iets om te noemen in de discussie**
According to a review of 705 published psychological meta-analyses (Van Erp et al., 2017),
these values of $\tau^2$ fall within the range observed in practice.
Note that both BRMA and RMA assume linear effects.
To test the robustness of the algorithms to violations of this assumption, true effect sizes were simulated using two models:
one with a linear effect of one moderator, $T_{i}= \beta x_{1i} + \epsilon_i$, and one with a non-linear (cubic) effect of one moderator, $T_{i}= \beta x_{1i} + \beta x_{1i}^{2} + \beta x_{1i}^{3} + \epsilon_i$,
where $\epsilon_i \sim N(0, \tau^2)$.
The algorithms further assume normality of residuals.
To examine robustness of the algorithms to violations of this assumption,
moderator variables were simulated as skewed normal moderators, with scale parameter $\omega \in (0, 2, 10)$, where $\omega = 0$ corresponds to the standard normal distribution.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 100 data sets were generated. 
In each data set, the observed effect size $y_i$ was simulated as a standardized mean difference (SMD),
sampled from a non-central t-distribution.

@@

BRMA aims to address the challenge that arises when meta-analysing heterogeneous bodies of literature, with few studies relative to the number of moderators.
BRMA can be used to identify relevant moderators when it is not known beforehand which moderators are responsible for between-studies differences in observed effect sizes.
To facilitate adoption of this method in applied research, we have published the function `brma()` in the R package `pema`.
Here, we offer several recommendations for its use.
The first recommendation precedes analysis, and relates to
the design of the meta-analysis.
When the search for moderators is exploratory,
researchers ought to be inclusive, but focus on moderators that are expected to be
relevant, including theoretically relevant moderators, as well as moderators pertaining
to the sample, methods, instruments, study quality, and publication type.
In our experience, many applied researchers code such study characteristics anyway, but omit
them from their analyses for lack of statistical power.
Moderators can be continuous or categorical, in which case they should be dummy-coded.
Missing data must be accounted for.
The best way to do so is by retrieving the missing information, by contacting authors or comparing different publications on the same data.
If missing data remains, users can either use a single imputation method (for example, a non-parametric imputation method like missForest),
or manually aggregate the results across multiple imputations.
The effect sizes and their variances must be computed using suitable methods;
note that many such methods are available in the R package `metafor` (Viechtbauer et al.,, 2010).
With regard to data analysis, we recommend the use of a horseshoe prior by default,
because it demonstrated the best predictive performance and most attractive trade-off between sensitivity and specificity in our simulations.

When estimating the model, it is important to ascertain that the algorithm has converged before interpreting the results.
Stan, the computational back-end of `brma()`, should return warnings and errors if there are indicationsof non-convergence.
<!-- Sara, wat kan men nog meer doen om convergence te checken? Iets met trace plots? Alle standaard plots voor stan-model objecten zijn beschikbaar, dankzij de functie `as.stan()`. Misschien kan je wat tips geven hier? -->

>>>>

$$
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
$$
The resulting parameter estimates are those that give the best predictions of the dependent variable in the present data set.
Penalized regression, by contrast, adds a penalty term to the RSS.
One commonly used penalty is the L1-norm of the regression coefficients, or LASSO penalty ([@hastieElementsStatisticalLearning2009]),
which corresponds to the sum of their absolute values.
This gives the penalized residual sum of squares:

@@

The standard linear model estimates an intercept, which reflects the expected value of the outcome when all predictors are equal to zero, and regression coefficients for the effect of moderators.
In some cases, it may be desirable to omit the intercept.
For example, if an analysis contains categorical predictors,
these can be encoded as dummy variables, with values $x \in \{0, 1\}$.
For a variable with $c$ categories, the number of dummy variables must be equal to $c-1$;
the omitted category functions as a reference category, and its expected value is represented by the model intercept $b_0$.
This so-called *regression specification* of a model may be useful when there is a meaningful reference category.
For example, imagine a study on the effectiveness of interventions for specific phobia with two types of interventions: An industry-standard intervention, and a more recently developed one.
In this case, it might make sense to code the industry standard intervention as the reference category,
and dummy-code the new contender.
The model will then estimate whether the newly developed intervention has an effect size significantly lower or higher than the industry standard.
In other cases, there may not be a straightforward reference category.
For example, imagine a study on the effectiveness of one intervention for specific phobia in two continents.
In such cases, the average effect in both continents may be estimated by omitting the intercept, and including all $c$ dummy variables.
This so-called *ANOVA specification* of a model estimates a mean for all dummy-coded categories.
In BRMA, as in other R functions, one can use ANOVA specification by explicitly removing the intercept from the model formula; for example, if `yi` is the effect size and $C$ a categorical moderator, regression specification with $c-1$ dummies is specified as `yi ~ C`, and ANOVA specification with $c$ dummies is specified as `yi ~ -1 + C`.
<!-- This part seems a repetition from the info on standardization of dummies above?: 
The appropriate standardization of dummy variables in penalized regression analyses is a topic of ongoing debate.
Suffice it to say that, for greater interpretability, users may want the freedom to leave dummy variables unstandardized.
This is possible in `brma()` by ... 
-->

@@

Many different prior distributions exist that have specific properties that enable them to shrink small regression coefficients towards zero, while keeping substantial coefficients large ([@vanErpOberskiMulder2019]). Some prior distributions have been shown to result in exactly the same solutions as classical, frequentist penalties, such as the LASSO. Specifically, placing independent double exponential (i.e., Laplace) priors on the regression coefficients can result in posterior modes that are equal to the classical LASSO estimates ([@ParkCasella2008]). In addition, many prior distributions have been developed and investigated specifically for the purpose of providing good shrinkage properties, meaning that the prior pulls small regression coefficients towards zero, without exerting any influence on substantial regression coefficients. A popular prior in this regard is the horseshoe prior ([@CarvalhoPolsonScott2010]). An advantage of the horseshoe prior compared to the lasso is the fact that it has heavier tails. As a result, it will not shrink and therefore bias substantial coefficients as much as the LASSO prior with its lighter tails.

@@

Bayesian penalized meta-analysis is implemented in the function `brma()` in the R-package `pema`.
For estimation, it depends on Stan, a probabilistic programming language that uses Hamiltonian Monte Carlo to sample from the posterior distribution ([@StanManual]).
Being written in C++, Stan is computationally efficient, but models must be compiled prior to estimation.
This results in substantial computational overhead.
To avoid this overhead, `pema` uses pre-compiled models corresponding to random-effects and three-level meta-regression, with and without an intercept.
Future updates may bring additional models.
At the time of writing, `brma()` supports two priors: the LASSO and the regularized horseshoe.
The LASSO prior is implemented as follows:
<!-- CJ: Sara, kan jij checken of La de juiste notatie is voor de Laplace prior? 
SvE: ik heb er DE van gemaakt voor double exponential, aangezien dat de aanduiding is die ook in Stan en de Stan manual wordt gebruikt.
-->

$$
\beta_{j} \sim \text{DE}(0, \frac{s}{\lambda})
$$
where DE denotes the double exponential distribution with a location equal to 0 and a scale determined by a global scale parameter $s$ and an inverse-tuning parameter $\lambda$.
By default in `brma()`, the global scale parameter is set to 1,
and the inverse-tuning parameter is given a $\chi^2$ prior with 1 degree of freedom.
Its value is thus optimized during model estimation.

@@

$$
\begin{aligned}
\beta_{j} &\sim N(0, \tilde{\tau}_j^2\lambda) \text{, with } \tilde{\tau}_j^2 = \frac{c^2 \tau_j^2}{c^2 + \lambda^2 \tau_j^2}\\
\lambda &\sim \text{student-}t^+(\nu_1, 0, \lambda^2_0)\\
\tau_j &\sim \text{student-}t^+(\nu_2, 0, 1)\\
c^2 &\sim \Gamma^{-1}(\frac{\nu_3}{2}, \frac{\nu_3 s^2}{2})
\end{aligned}
$$
where N denotes the normal distribution, student-$t^+$ denotes the half-t distribution and $\Gamma^{-1}$ denotes the inverse Gamma distribution. This extension of the horseshoe is more numerically stable in certain cases. <!-- CJ: Je spreekt hier over het argument par_ratio van de functie in Stan. Maar hier in de tekst verwacht ik uitleg over de formule. Zien we par_ratio terug in die formule?; SvE: ik heb dit hieronder nu extra uitgelegd -->
In this formula, $\lambda^2_0$ is a global scale parameter that affects the overall shrinkage of the prior, with smaller values resulting in more shrinkage. The default value in `brma()` is 1. However, if prior information regarding the number of relevant moderators is available, it is best to include this information by setting $\lambda^2_0 = \frac{p_0}{p-p_0}{\frac{\sigma}{\sqrt{n}}$, where $p_0$ represents the number of relevant moderators, $p$ the total number of moderators, $\sigma$ is the residual standard deviation and $n$ equals the number of observations. This can be done by setting the argument `relevant_pars` equal to the expected number of relevant moderators. 
<!-- SvE: nu twijfel ik toch weer of dit helemaal goed gaat. sigma is in ons model namelijk op 0 gezet en wordt hier dus niet in meegenomen. Klopt het dat de residual SD 0 moet zijn in dit model? De rest van de global scale wordt berekend buiten het Stan model om in de brma functie zelf. -->
The thickness of the tails is controlled by two degrees of freedom parameters, $\nu_1$ and $\nu_2$, which default to 1 in `brma()`.
Increasing these degrees of freedom parameters results in a prior with lighter tails,
which is strictly no longer a horseshoe prior.
However, in cases where the model is weakly identified, for example when there are more moderators than observations, these lighter tails can aid model convergence. 
The regularized horseshoe differs from the standard horseshoe in the specification of a finite "slab". This slab ensures at least some regularization of large coefficients and as a consequence,
more stable results.
This slab is governed by a degrees of freedom parameter ($\nu_3$, set to 4) and a scale parameter ($s$, set to 1). <!--Sara, eigenlijk is het niet helemaal duidelijk wat al deze nummers doen. Kunnen we een interactieve plot maken waarmee gebruikers kunnen zien hoe aanpassingen de vorm van de prior beinvloeden?; SvE: ik heb een prototype shiny app gemaakt.-->

@@

Predictive performance reflects how well the algorithm is able to predict data not used to estimate the model parameters, in other words, it indicates the generalizability of the model.
To compute it, for each iteration of the simulation both a training dataset and a testing dataset are generated.
The model is estimated on the training data, which has a varying number of cases according to the simulation conditions.
Predictive performance is then operationalized as the explained variance in the testing data, $R^2_{test}$.
The testing data has 100 cases in all simulation conditions.
The $R^2_{test}$ reflects the fraction of variance in the testing data explained by the model,
relative to the variance explained by the mean of the training data.
For a predictive performance measure, it is necessary to use the mean of the training data, as the mean of the testing set is a descriptive statistic of that sample. **ik vind deze zin niet heel duidelijk** <!-- SvE: dit was nog een eerdere comment van mij (toen ik nog niet wist hoe ik echt kon commenten binnen Rmarkdown :) ) -->
The resulting metric $R_{test}^{2}$ is expressed by the following equation:

@@

To examine performance in a range of realistic meta-analysis scenarios,
several design factors were manipulated:
The number of studies in the training data $k \in (20, 40, 100)$, the average within-study sample size $\bar{n} \in (40, 80, 160)$, the population effect size of relevant moderators $\beta \in (0, .2, .5, .8)$, the number of moderators $p \in (2, 3, 6)$, and residual heterogeneity $\tau^{2} \in (.01, .04, .1)$. **het aantal moderatoren is redelijk laag; ik zou vooral nog voordelen van brma verwachten met meer moderatoren maar is dat realistisch in de praktijk? Zo ja, dan is het iets om te noemen in de discussie** <!-- hier ook nog een eerdere comment. Mocht je een keer een toepassing tegenkomen met echt veel moderatoren, kan het interessant zijn hier nog een kleine simulatie op te baseren. Een voorbeeld kan zijn een set van moderatoren die de kwaliteit van studies proberen te vatten zoals we laatst hebben besproken -->
According to a review of 705 published psychological meta-analyses (Van Erp et al., 2017),
these values of $\tau^2$ fall within the range observed in practice.
Note that both BRMA and RMA assume linear effects.
To test the robustness of the algorithms to violations of this assumption, true effect sizes were simulated using two models:
one with a linear effect of one moderator, $T_{i}= \beta x_{1i} + \epsilon_i$, and one with a non-linear (cubic) effect of one moderator, $T_{i}= \beta x_{1i} + \beta x_{1i}^{2} + \beta x_{1i}^{3} + \epsilon_i$,
where $\epsilon_i \sim N(0, \tau^2)$.
The algorithms further assume normality of residuals.
To examine robustness of the algorithms to violations of this assumption,
moderator variables were simulated as skewed normal moderators, with scale parameter $\omega \in (0, 2, 10)$, where $\omega = 0$ corresponds to the standard normal distribution.
The design factors combined to produce `r out$conditions` unique conditions.
For all simulation conditions, 100 data sets were generated. 
In each data set, the observed effect size $y_i$ was simulated as a standardized mean difference (SMD),
sampled from a non-central t-distribution.

@@

BRMA aims to address the challenge that arises when meta-analysing heterogeneous bodies of literature, with few studies relative to the number of moderators.
BRMA can be used to identify relevant moderators when it is not known beforehand which moderators are responsible for between-studies differences in observed effect sizes.
To facilitate adoption of this method in applied research, we have published the function `brma()` in the R package `pema`.
Here, we offer several recommendations for its use.
The first recommendation precedes analysis, and relates to
the design of the meta-analysis.
When the search for moderators is exploratory,
researchers ought to be inclusive, but focus on moderators that are expected to be
relevant, including theoretically relevant moderators, as well as moderators pertaining
to the sample, methods, instruments, study quality, and publication type.
In our experience, many applied researchers code such study characteristics anyway, but omit
them from their analyses for lack of statistical power.
Moderators can be continuous or categorical, in which case they should be dummy-coded.
Missing data must be accounted for.
The best way to do so is by retrieving the missing information, by contacting authors or comparing different publications on the same data.
If missing data remains, users can either use a single imputation method (for example, a non-parametric imputation method like missForest),
or manually aggregate the results across multiple imputations.
The effect sizes and their variances must be computed using suitable methods;
note that many such methods are available in the R package `metafor` (Viechtbauer et al., 2010).
With regard to data analysis, we recommend the use of a horseshoe prior by default,
because it demonstrated the best predictive performance and most attractive trade-off between sensitivity and specificity in our simulations.

When estimating the model, it is important to ascertain that the algorithm has converged before interpreting the results.
Stan, the computational back-end of `brma()`, should return warnings and errors if there are indications of non-convergence.
<!-- Sara, wat kan men nog meer doen om convergence te checken? Iets met trace plots? Alle standaard plots voor stan-model objecten zijn beschikbaar, dankzij de functie `as.stan()`. Misschien kan je wat tips geven hier?; SvE: ik zal dit verwerken in de toepassing. -->
